### 2009-12-18 doc/tech_doc/a-rex/arex_tech_doc.tex
Small documentation and example update to reflect the configuration of transfer shares through XML
add new cache features

The A-REX can cache input files, so that subsequent jobs requiring the
same files do not have to download them again. Caching is enabled if
one or more cache directories are specified in the configuration
file. All input files except files uploaded by the user during job
submission are cached by default. This includes excutable files
downloaded by the A-REX. Caching can be explicitly turned off by the
user in the job description (see~\cite{userguide1}). The disk space
occupied by the cache is controlled by removing files in the order of
least recent access. For more information on configuration see
Section~\ref{SubSection:ConfigFile}.

@@

Cached files are stored in sub-direcories under the \emph{data} directory
in each main cache directory. Filenames are constructed from an SHA-1
hash of the URL of the file, and split into subdirectories based on
the two initial characters of the hash. This enables the cached files
to be evenly split over a number of subdirectories. If
more than one cache directory is used the initial letter of the hash
also determines which cache is to be used. The algorithm is simply
to use the cache directory at index \emph{i} in the list of directories,
where \emph{i} is found from the initial letter of the hash mod the
number of caches. This limits the maximum number of cache directories
to 16, as SHA-1 hashes use the character set {[}\emph{0-9,a-f}]. In
the extremely unlikely event of a collision between two URLs having
the same SHA-1 hash, caching will not be used for the second file.

There is no indexing system for the cache, and any cache filename
can be easily determined from a URL by using the same hashing algorithm
as the A-REX, e.g. the standard command line tool \emph{sha1sum. }The
cache directory used by a file can be determined from the initial
letter of the hash and the order of cache directories in the configuration
file. Some associated metadata (the corresponding URL and an expiry
time, if available) are stored in a file with the same name as the
cache file, with a \emph{.meta} suffix.

@@

If a job requests an input file which can be cached or is allowed to
be cached, it is stored in the selected cache directory, and depending
on the configuration, either the file is copied to the SD or a hard
link is created in a per-job directory and a soft link is created in
the SD to there. The per-job directories are in the \emph{joblinks}
subdirectory of the main cache directory. The former option is advised
if the cache is on a file system which will suffer poor performance
from a large number of jobs reading files on it, or the file system
containing the cache is not accessible from worker nodes. The latter
option is the default option. The per-job directory is only readable
by the local user running the job, and the cache directory is readable
only by the A-REX user. This means that the local user cannot access
any other users' cache files. It also means that cache files can be
removed without needing to know whether they are in use by a currently
running job. Note that when running A-REX under a non-privileged user
account, all cache files will be owned and accessible by the same
user, and therefore modifyable by running jobs. This is potentially
dangerous and so cacheing should be used with caution in this case.

@@

@@

@@

\begin{itemize}
\item \textbf{\textit{mail}}\textit{=e-mail\_address} -- specifies an email
address \textbf{from} which notification mails are sent.
\item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} -- specifies the
time in seconds for the SD to be available after job finishes (\emph{ttl})
and after job was deleted (\emph{ttr}) due to \emph{ttl}. Defaults
are 7 days for \emph{ttl} and 30 days for \emph{ttr}. The minumum value
for both parameters is 2 hours.
\item \textbf{\textit{lrms}}\textit{=default\_lrms\_name default\_queue\_name}
-- specifies names for the LRMS and queue. Queue name can also be specified
in the JD (currently it is not allowed to override LRMS by using
the JD).
\item \textbf{\textit{sessiondir}}\textit{=path} -- specifies path to
  the directory in which the SD is created. If \emph{path} is {*} the
  default sessiondir is used -- \textit{\$HOME/.jobs}.
\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} -- specifies
a directory to store cached data (see section \ref{sec:cache}). Multiple
cache directories may be specified by specifying multiple \emph{cachedir}
commands. Cached files will be distributed evenly between the caches.
Specifying no \emph{cachedir }command or commands with an empty path
disables caching. Optional \textit{link\_path} specifies the path
at which \emph{path} is accessible on computing nodes, if it is different
from the path on the A-REX host. If \textit{link\_path} is set to '.'
files are not soft-linked, nor are per-job links created, but files
are copied to the session directory.
\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} --
  specifies high and low watermarks for space used on the file system
  on which the cache directory is located, as a percentage of total
  file system capacity. When the max is exceeded, files will be
  deleted to bring the used space down to the min level. It is a good
  idea to have each cache on its own separate file system. To turn off
  this feature, \char`\"{}cachesize\char`\"{} without parameters can
  be specified. These cache settings apply to all caches specified by
  \emph{cachedir} commands.
\item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number
of times job will be allowed to rerun after it failed at any stage.
Default value is \emph{5}. This only specifies a upper limit. The actual
number is provided in job description and defaults to 0.
\end{itemize}

@@

>>>>

The A-REX can cache input files, so that subsequent jobs requiring the
same files do not have to download them again. Caching is enabled if
one or more cache directories are specified in the configuration
file. All input files except files uploaded by the user during job
submission are cached by default. This includes executable files
downloaded by the A-REX. Caching can be explicitly turned off by the
user in the job description (see~\cite{userguide1}). The disk space
occupied by the cache is controlled by removing files in the order of
least recent access. For more information on configuration see
Section~\ref{SubSection:ConfigFile}.

@@

Cached files are stored in sub-directories under the \emph{data}
directory in each main cache directory. Filenames are constructed from
an SHA-1 hash of the URL of the file and split into subdirectories
based on the two initial characters of the hash. In the extremely
unlikely event of a collision between two URLs having the same SHA-1
hash, caching will not be used for the second file.

When multiple caches are used, a new cache file goes to a randomly
selected cache, where each cache is weighted according to the size of
the file system on which it is located.  For example: if there are two
caches of 1TB and 9TB then on average 10\% of input files will go to
the first cache and 90\% will go to the second cache.

Some associated metadata including the corresponding URL and an expiry
time, if available, are stored in a file with the same name as the
cache file, with a \emph{.meta} suffix.

@@

If a job requests an input file which can be cached or is allowed to
be cached, it is stored in the selected cache directory, and depending
on the configuration, either the file is copied to the SD or a hard
link is created in a per-job directory and a soft link is created in
the SD to there. The per-job directories are in the \emph{joblinks}
subdirectory of the main cache directory. The former option is advised
if the cache is on a file system which will suffer poor performance
from a large number of jobs reading files on it, or the file system
containing the cache is not accessible from worker nodes. The latter
option is the default option.  Files marked as executable in the job
will be stored in the cache without executable permissions, but they
will be copied to the SD and the appropriate permissions applied to
the copy.

The per-job directory is only readable by the local user running the
job, and the cache directory is readable only by the A-REX user. This
means that the local user cannot access any other users' cache
files. It also means that cache files can be removed without needing
to know whether they are in use by a currently running
job. \textbf{IMPORTANT:} If a cache is mounted from an NFS server and
the GM is run by the root user, the server must have the
\emph{no\_root\_squash} option set for the GM host in the
\emph{/etc/exports} file, otherwise the GM will not be able to create
the required directories. Note that when running A-REX under a
non-privileged user account, all cache files will be owned and
accessible by the same user, and therefore modifyable by running
jobs. This is potentially dangerous and so cacheing should be used
with caution in this case.

@@

\subsection{Remote Caches}
\label{sec:remotecaches}

If a site has multiple A-REXs running, an A-REX can be configured to
have its own caches and have read-only access to caches under the
control of other A-REXs (remote caches). An efficient way to reduce
network traffic within a site is to configure A-REXs to be under
control of caches on their local disks and have caches on other hosts
as remote caches. If an A-REX wishes to cache a file and it is not
available on the local cache, it searches for the file in remote
caches. If the file is found in a remote cache, the actions the A-REX
takes depends on the policy for the remote cache. The file may be
replicated to the local cache to decrease the load on the remote file
system caused by many jobs accessing the file. However, this will
decrease the total number of cache files that can be stored. The other
policy is to use the file in the remote cache, creating a per-job
directory for the hard link in the remote cache. Then the link is
created from the session dir to that directory, bypassing the local
cache completely. The usual permission and validity checks are
performed for the remote file. Note that no creation or deletion of
remote cache data is done - cache cleaning is only performed on local
caches.

@@

Caches can be added to and removed from the configuration as required
without affecting any cached data, but after changing the
configuration file, the A-REX should be restarted. If a cache is to be
removed and all data erased, it is recommended that the cache be put
in a \emph{draining} state until all currently running jobs possibly
accessing files in this cache have finished. In this state the cache
will not be used by any new jobs, but the hard links in the
\emph{joblinks} directory will be cleaned up as each job
finishes. Once this directory is empty it is safe to delete the entire
cache. See the \emph{cachedir} option in
Section~\ref{SubSection: ConfigFile} for how to set a cache to a
draining state.

@@

\begin{itemize}
\item \textbf{\textit{mail}}\textit{=e-mail\_address} -- specifies an email
address \textbf{from} which notification mails are sent.
\item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} -- specifies the
time in seconds for the SD to be available after job finishes (\emph{ttl})
and after job was deleted (\emph{ttr}) due to \emph{ttl}. Defaults
are 7 days for \emph{ttl} and 30 days for \emph{ttr}. The minumum value
for both parameters is 2 hours.
\item \textbf{\textit{lrms}}\textit{=default\_lrms\_name default\_queue\_name}
-- specifies names for the LRMS and queue. Queue name can also be specified
in the JD (currently it is not allowed to override LRMS by using
the JD).
\item \textbf{\textit{sessiondir}}\textit{=path} -- specifies path to
  the directory in which the SD is created. If \emph{path} is {*} the
  default sessiondir is used -- \textit{\$HOME/.jobs}.
\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} --
  specifies a directory to store cached data (see section
  \ref{sec:cache}). Multiple cache directories may be specified by
  specifying multiple \emph{cachedir} commands. Cached data will be
  distributed over multiple caches according to free space in
  each. Specifying no \emph{cachedir} command or commands with an
  empty path disables caching. The optional \textit{link\_path} specifies
  the path at which \emph{path} is accessible on computing nodes, if
  it is different from the path on the A-REX host. If
  \textit{link\_path} is set to '.'  files are not soft-linked, nor
  are per-job links created, but files are copied to the session
  directory. If a cache directory needs to be drained, then
  \emph{cachedir} should specify \emph{``drain''} as the
  \emph{link\_path}.
\item \textbf{\textit{remotecachedir}}\textit{=path {[}link\_path]} -
  specifies caches which are under the control of other A-REXs, but which
  this A-REX can have read-only access to (see
  Section~\ref{sec:remotecaches}).  Multiple remote cache directories may be
  specified by specifying multiple \emph{remotecachedir} commands. If
  a file is not available in paths specified by cachedir, the GM looks
  in remote caches. \emph{link\_path} has the same meaning as in
  \emph{cachedir}, but the special path ``replicate'' means files will
  be replicated from remote caches to local caches when they are requested.
\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} --
  specifies high and low watermarks for space used on the file system
  on which the cache directory is located, as a percentage of total
  file system capacity. When the max is exceeded, files will be
  deleted to bring the used space down to the min level. It is a good
  idea to have each cache on its own separate file system. To turn off
  this feature, \char`\"{}cachesize\char`\"{} without parameters can
  be specified. These cache settings apply to all caches specified by
  \emph{cachedir} commands.
\item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number
of times job will be allowed to rerun after it failed at any stage.
Default value is \emph{5}. This only specifies a upper limit. The actual
number is provided in job description and defaults to 0.
\end{itemize}

@@

When XML file only is used to configure the A-REX, the transfer shares
can be implemented by defining \emph{maxLoadShare} (the limit itself)
and \emph{loadShareType} (the scheme used) elements inside 
\emph{loadLimits} block.
